A distributed key-value store using Raft

### 2014-06-04 proj/paper.tex
paper progress

Finally, AppendEntries messages also serve to keep the leader in power. Whenever a node receives an AppendEntries message with a term at least as high as its own, it will revert back to follower status and reset its timeout. Thus, if a leader sends out AppendEntries messages (possibly with zero actual entries attached) with greater frequency than once every $150$ms, no other nodes will attempt to become leader.
\section{Description of the implementation}
\subsection{Structure and requests}
The actual implementation is done in Python building on the provided example node as a base. There are two distinct components to this implementation -- the client node and the server node. The server node code implements the Raft state and its transitions and stores the replicated state of the cluster. The client node serves as a middleman between the cluster and the broker due to the broker's limited abilities to address nodes. A typical run with the broker scripts will start the client node

When a 
\end{document}

>>>>

Finally, AppendEntries messages also serve to keep the leader in power. Whenever a node receives an AppendEntries message with a term at least as high as its own, it will revert back to follower status and reset its timeout. Thus, if a leader sends out AppendEntries messages (possibly with zero actual entries attached) with greater frequency than once every $150$ms, no other nodes will attempt to become leader.
\section{Description of the implementation}
\subsection{Structure and requests}
The actual implementation is done in Python building on the provided example node as a base. There are two distinct components to this implementation -- the client node and the server node. The server node code implements the Raft state and its transitions and stores the replicated state of the cluster. The client node serves as a middleman between the cluster and the broker due to the broker's limited abilities to address nodes. A typical run with the broker scripts will start the client nodes first, then the server nodes.

\subsection{GET requests}
When the broker issues a get request, it should always be directed towards one of the clients. Upon receiving this request, the client node associates the info in the request to the id provided by the broker so that it can formulate a response back to the broker after the Raft cluster responds. It then generates a new id which is unique with respect to all other clients and requests by prepending its name to a counter incremented every time an id is generated. The client node will then associate this new id with the broker's id for the request, then send off a \verb|GET| message to who it thinks the Raft leader is.

The client node initially assumes that the first entry in its list of Raft nodes is the leader. However, if it sends a \verb|GET| message to a node which is not the leader, that node will reply with a \verb|redirect| message providing who it thinks that the current leader is. The client will then record this as the new leader and retry the request. If the cluster is currently transitioning between leaders, the response to the \verb|redirect| may be null. In this case, the client retries with a randomly chosen node.

Due to the nature of log replication, the leader will always have the most up-to-date replicated state if it is still the sole leader and has not been deposed. Thus, there is no need for the leader to append an entry into its log -- it can just respond using its copy of the replicated state after verifying that it has not been deposed. To do this, after receiving the \verb|GET| request from the client, the cluster leader sends out a possibly empty AppendEntries message tagged with the client's request id to all followers in order to solicit a response. It then associates the request id to client request, then exits to process other messages. The leader will keep track of the responses associated with that id, and once it receives a majority affirming that the term has not ended, it will reply back to the client with the appropriate data indicating success, including the client's request id that was sent with the initial request. When the client receives this message, it will retrieve the broker's request using the id in the reply, then formulate a reply back to the broker indicating success and carrying the value.

However, if a fault occurs during this process which leaves the leader unable to contact a majority of nodes, no response may be generated. To handle this case, the client nodes associate a $300$ms timeout with each request, and will retry the request if this timeout expires. The value of the timeout was chosen to be small while still being long enough to wait out leadership changes. After some number of retries (I picked $5$ to allow for two cycles of timeout-redirects), the client will return an error to the broker. 
\subsection{SET requests}
The beginning and end of a \verb|SET| request look much like those of \verb|GET| requests, wherein the broker submits a request to the client, the client repackages it for the cluster, and the response follows the same chain backwards. The middle part of a \verb|SET| request differs, as setting a value requires a change to be committed into the replicated state. To do this, the leader appends an entry into its log containing the key, the value, the request id, and the client's name. It then runs through the Raft consensus algorithm until the entry is committed. In this implementation, committing a \verb|SET| request entry on any node involves writing it to the replicated state, then messaging the client recorded in the entry to let it know that the request has been completed. This results in all nodes sending the same success message to the client, which improves the odds that the client will see one of them in the event of a network partition. However, it also means that the client has to deal with filtering out duplicates. It does this by using the unique request id, which is included in all messages sent from and to the client.
\end{document}
